Cloud correction and filtering of operational microwave humidity radiances with case specific uncertainty estimation

### 2020-09-18 Article/ici_submm_cloudcorr.tex
Minor work on Introduction.

\introduction
%
Satellite observations of humidity inside the troposphere are mainly performed
by downward-looking sensors. Among this class of observations, the frequency
range around 183\,GHz has a special position. Water vapour has a noticeable
transition at 22\,GHz, but it is relatively weak and only column values can be
derived \citep[e.g.][]{schluessel1990atmospheric} for the observation geometry
of concern. The first transition in the microwave region that can be used to
derive altitude information, i.e.\ can be used for ``sounding'', is the one at
183.31\,GHz \citep{kakar1983retrieval,wang1983profiling}. On the other hand, at
infrared wavelengths a high number of water vapour transitions are found,
including some of high strength. As a consequence, infrared sounders can
provide humidity profiles with high precision and good vertical resolution, but
with strong limitations imposed by clouds. To be able to also sense humidity
inside and below clouds weather satellites are since some time equipped with
channels around 183\,GHz. Today such channels are part of several sensors, such
as ATMS (Advanced Technology Microwave Sounder, \citet{weng2012introduction}).

@@

The main motivation for introducing 183\,GHz channels in operational sensors is
numerical weather prediction (NWP). Usage of microwave data by ``all-sky''
assimilation is growing \citep{geer2017growing}, but 183\,GHz data are still
mainly used in a clear sky fashion \citep{geer2018all}. The later is
particularly true in NWP of regional scope, all-sky assimilation of 183\,GHz at
many national weather agencies will likely not be reached in many years. The
cloud filtering in NWP differs between centres. One approach is to make use of
a ``scattering index'' \citep{bennartz2002precipitation}, based on the
observations alone. Today, likely more common is to use ``observation minus
background'' (O-B), where the forecast model is used to obtain an estimate of
the expected clear-sky value and the observation is rejected if the deviation
exceeds some threshold (???). In both cases, the filtering typically involves
observations around 89 and/or 150\,GHz.

Despite the broad usage these filtering methods have some drawbacks. For
183\,GHz, the impact of hydrometeors systematically causes a decrease in the
observed radiance (maybe with exceptions at extremely dry conditions), see
e.g.\ \citet{barlakas:three:20} and below. This means that if any cloud
contamination is missed by the filtering this will cause a negative bias in
mean radiance, compared to the true clear-sky mean. This bias will translate to
a bias in humidity after the retrieval or assimilation. The alternative is to
apply a very strict filtering, but this will result in that a high fraction of
actually clear-sky values will be rejected \todo{(Reka or better reference)},
i.e.\ an important loss of useful data.

@@

A maybe less obvious problem is the uncertainty to assign to the filtered
values. To our best knowledge, in best case estimates of mean and worst case
errors are provided. Some cases with relatively high cloud impact will likely be
missed, while most cases are clear sky from start. As the remaining cloudy
cases can cause significant biases, the likely solution is to apply a quite
conservative (high) error estimate. However, this will unnecessary downgrade
the value of the truly clear sky cases and the observations are used in a
non-optimal manner.

@@

The estimation of NFSC values makes use of a special type of machine learning,
denoted as Quantile Regression Neural Network (QRNN,
\citet{pfreundschuh:aneur:18}). Unlike standard usage of machine learning where
only some kind of best estimate is provided, QRNN works in a Bayesian fashion
and instead gives a description of the posterior uncertainty. More precisely,
QRNN outputs an user specified set of percentiles that gives a discrete
description posterior distribution, that can be process to derive e.g.\ the
expectation value.

>>>>

\introduction
%
Satellite observations of humidity inside the troposphere are mainly performed
by downward-looking sensors. Among this class of observations, the frequency
range around 183\,GHz has a special position. Water vapour has a noticeable
transition at 22\,GHz, but it is relatively weak and only column values can be
derived \citep[e.g.][]{schluessel1990atmospheric} for the observation geometry
of concern. The first transition in the microwave region that can be used to
derive altitude information, i.e.\ can be used for ``sounding'', is the one at
183.31\,GHz \citep{kakar1983retrieval,wang1983profiling}. On the other hand, at
infrared wavelengths a high number of water vapour transitions are found,
including some of high strength. As a consequence, infrared sounders can
provide humidity profiles with high precision and good vertical resolution, but
with strong limitations imposed by clouds. To be able to also sense humidity
inside and below clouds, weather satellites are since some time equipped with
channels around 183\,GHz. Today such channels are part of several sensors, such
as ATMS (Advanced Technology Microwave Sounder, \citet{weng2012introduction}).

@@

The main motivation for introducing 183\,GHz channels in operational sensors is
numerical weather prediction (NWP). Usage of passive microwave data by
``all-sky'' assimilation in global NWP is growing in \citep{geer2017growing},
but 183\,GHz data are still mainly used in a clear sky fashion
\citep{geer2018all}. The later is particularly true in NWP of regional scope
\citep{gustafsson2018survey}, all-sky assimilation of 183\,GHz at many national
weather agencies will likely not be reached in many years. Anyhow, the cloud
filtering applied differs between NWP centres. One approach is to make use of a
``scattering index'' \citep{bennartz2002precipitation}, based on the
observations alone. Today, likely more common is to use ``observation minus
background'' (O-B), where the forecast model is used to obtain an estimate of
the expected clear-sky value and the observation is rejected if the deviation
exceeds some threshold (???). In both cases, the filtering typically involves
observations around 89 and/or 150\,GHz.

Despite the broad usage these filtering methods have some drawbacks. For
183\,GHz, the impact of hydrometeors systematically causes a decrease in the
observed radiance (maybe with exceptions at extremely dry conditions), see
e.g.\ \citet{barlakas:three:20} and below. This means that if any cloud
contamination is missed by the filtering this will cause a negative bias in
mean radiance, compared to the true clear-sky mean. This bias will translate to
a bias in humidity after the retrieval or assimilation. The alternative is to
apply a very strict filtering, but this will result in that a high fraction of
actually clear-sky values will be rejected, i.e.\ an important loss of useful
data.

@@

A maybe less obvious problem is the uncertainty to assign to the filtered
values. To our best knowledge, so far only estimates of mean and worst case
errors have been provided. Some cases with relatively high cloud impact will
likely be missed, while most cases are clear sky from start. As the remaining
cloudy cases can cause significant biases, the likely solution is to apply a
quite conservative (high) error estimate. However, this will unnecessary
downgrade the value of the truly clear sky cases and the observations are used
in a non-optimal manner.

@@

The estimation of NFCS values makes use of a special type of machine learning,
denoted as Quantile Regression Neural Network (QRNN,
\citet{pfreundschuh:aneur:18}). Unlike standard usage of machine learning where
only some kind of best estimate is provided, QRNN works in a Bayesian fashion
and instead gives a description of the posterior uncertainty. More precisely,
QRNN outputs an user specified set of percentiles that gives a discrete
description posterior distribution, that can be process to derive e.g.\ the
expectation value.
